### 2022-11-07 Manuscript/manuscript.tex
add sections

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[margin=1 in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{lipsum}
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

\title{The Effect of The COVID-19 Pandemic on Influenza Vaccination Rates in the United States}
\author{Jessica Zambuto\\
Department of Statistics, University of Connecticut}

\begin{document}
\maketitle
\section*{Abstract}
\label{sec:Abstract}
\section*{Keywords}
\label{sec:Keywords}
\section{Introduction}
\label{sec:Introduction}
\section{Data}
\label{sec:data}
\section{Methods}
\label{sec:Methods}
\section{Results}
\label{sec:Results}
\section{Discussion}
\label{sec:Discussion}
\bibliographystyle{chicago}
\bibliography{citations}
\end{document}